---
title: "Assn5"
author: "Emily Bierbaum"
date: "2022-10-20"
output: html_document
---
## **Assignment 5**

#### Examine the R script for comments  
[R Script](Assn_5.R)  

##### Download the required R packages for this assignment
```{r}
require(tidyverse)
require(readr)
require(dplyr)
```

##### Import the part 1 and 2 files from GitHub 
```{r}
Part1_Assn5<-read_csv("https://github.com/mbtoomey/Biol_7263/blob/main/Data/assignment6part1.csv?raw=true",show_col_types = FALSE) 

Part2_Assn5<-read_csv("https://github.com/mbtoomey/Biol_7263/blob/main/Data/assignment6part2.csv?raw=true", show_col_types = FALSE)
```


##### A new file is created "Part1_tibble" 
```{r}
Part1_tibble <- Part1_Assn5 %>% pivot_longer(cols = starts_with("Sample"),
                                             names_to = c("Sample", "Sex", "Group"),
                                             names_prefix = "Sample", names_transform = list(Sample = as.integer),
                                             names_sep = "_") %>%
  pivot_wider(names_from = ID, values_from = value)
```
  
##### A new file is created "Part2_tibble"
```{r}
Part2_tibble <- Part2_Assn5 %>% pivot_longer(cols = starts_with("Sample"), 
                                             names_to = c("SampleTreatment"), 
                                             names_prefix = "Sample",
                                             names_transform = list(Sample = as.integer), 
                                             values_to = "count") %>% separate(SampleTreatment,
                                             into = c("Sample", "Treatment"), 
                                             convert = TRUE) %>% 
  pivot_wider(names_from = ID, values_from = count) %>% select(-Treatment)
```

##### Combine the two tibbles (part 1 and 2), and export to a csv file
```{r}
Part1_tibble %>% full_join(Part2_tibble, by = "Sample") -> Combined_Data

write_csv(Combined_Data, "Results/Combined_Assn5_Data.csv")
```

##### Create the variable residual mass and calculate the mean and standard deviation for group type and sex
```{r}
Combined_Data %>% transmute(Sex=Sex, Group=Group, resid_mass=mass/body_length) %>% 
  group_by(Group, Sex) %>% 
  summarize(mean_mass = mean(resid_mass, na.rm = TRUE), SD_mass = sd(resid_mass, na.rm = TRUE)) -> Residual_Mass
```

##### Export csv file
```{r}
write_csv(Residual_Mass, "Results/Final_Residual_Mass.csv")
```

[Pivot & Merge Data](Results/Combined_Assn5_Data.csv)  

[Residual Mass Statistics](Results/Final_Residual_Mass.csv)