---
title: "Assn_6"
author: "Emily Bierbaum"
date: "2022-11-11"
output: html_document
---

## **Assignment #6**

>

#### Load the required libraries
```{r}
require(tidyverse)
```

>
 
 
#### Import the csv file 
```{r}
MBT_ebird<-read.csv("Data/MBT_ebird2.csv")
```
>

#### **(1a)** Using the ebird dataset, calculate the number of species seen each month of each year in each location.
  - Group the csv files by location, month, and year  
  - Then we will summarize the groups by species count
  - `Length ()` function calculates the length of lists/vectors that have *unique* common names
  - `.group="drop"` is used because the output becomes ungrouped
```{r}
bird_stats<- MBT_ebird %>%
  group_by(location, month, year) %>%
  summarize(species_count = length(unique(common_name)), .groups = "drop")
```
 
>

#### **(1b)** Plot the number of species seen each month with the color of the points indicating year and facet this plot by location
  - Create a `ggplot` using the data above, `bird_stats`
  - `aes` specifies the variables to be plotted
  - `as.factor` is needed or else the months are in increments of 0.5
  - `facet_wrap` generates multiple plots based on location
  - The geometric is drawn in points and the axes and title were labeled
  - The figure legend will be for year
  - Lastly, I changed the color gradient to rainbow

```{r}
prob1 <- ggplot(data = bird_stats)+
  aes(as.factor(month), species_count, color = year) +
  facet_wrap(~location) +
  geom_point(size=1.5) +
  theme_gray() +
  xlab("Month") +
  ylab("Species Count") +
  ggtitle("Species Count by State") 

problem1<-prob1+scale_color_gradientn(colours = rainbow(5))
problem1
```

  

#### *Using the data set from Assignment #5:*
#### **(2)** Plot a comparison of mass by treatment including the individual observations, the mean, and standard error of the mean. Use point color or shape to indicate the sex.  
  - Import the data file from Assignment 5
  - Create a `ggplot` of the data frame `mass_treat`
  - Group and mass were specified as the variables to be plotted
  - `geom_jitter` adds a small amount of random variation for each point
  - The axes and title was labeled, as well as a cross bar that is the mean (purple), and error bars in black
  - The figure legend will be for sex
  - Error message is for the removed NA values

```{r}
mass_treat<-read.csv("Results/Combined_Assn5_Data.csv")

prob2 <-ggplot(data= mass_treat,aes(Group,mass)) +
  geom_jitter(size = 2, aes(Group, mass, color=Sex)) +
  xlab("Group") +
  ylab("Mass") +
  stat_summary(fun = mean,  
               geom = "crossbar", 
               width = 0.5, 
               color = "purple") +
  stat_summary(geom = "errorbar",  
               width = 0.6)+
  ggtitle("Mass by Group and Sex") 
  labs(color="Sex")
prob2
```


#### **(3)** Generate a scatter plot of age and mass, indicate treatment with point shape or color, and fit separate regression lines (without CI) to each treatment.  
  - Create a `ggplot` of the data frame with age and mass as the variables
  - Want geometric plot as points and create a figure legend for group (dots)
  - Label the axes and the title
  - `geom_smooth` adds a trend line and `method=lm` plots a linear model (linear regression line)
  - `se=false` removes the plotting of a confidence interval around the line
  - The `labs()` function sets the group (control vs treatment) as the figure legend for the linear model

```{r}
prob3<-ggplot(data= mass_treat,aes(age,mass) )+ 
  geom_point(size =3, aes(age, mass, color=Group))+ 
  xlab("Age")+ 
  ylab("Mass")+
  geom_smooth(size = 2, method = lm, 
              aes(color = Group,  group = Group), 
              se=FALSE)+
  labs(color="Group") + 
  ggtitle("Mass by Age and Group")

prob3
```


#### **(4)** Combine the plots from question 2 and 3 using patchwork tag each panel with, and number or letter and include a title for the overall plot.       
  - Download the library `patchwork`
  - `plot_annotation` adds the two plots together and adds a title over them both
  - The plots are then labeled A and B 

```{r}
require(patchwork)

prob2+prob3+plot_annotation(title = "Mass by Group and Age",
                              tag_levels = "A")

```




